---
title: "Assignment 2 - Final"
author: "Gianina Yumul"
date: "9/22/2020"
output: html_document
html_document:
    theme: cosmo
    toc: true
    toc_depth: 3
    toc_float: true
---
```{r}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(rgeos)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(readr)

#Assignment description: Create a map of the area, with its associated subareas, using each of three different map projections. Indicate the map projection you've used for each. Identify a variable and use at least three different methods to show how it varies along the subareas.

```


```{r}
US_states <- ne_states(country = "United States of America", returnclass = "sf") %>%
  filter(name != "Alaska",
         name != "Hawaii")

US_map <- ggplot(US_states) +
  geom_sf() +
  theme_bw()
```

Added the following projections (source = spatialreference.org):

"USGS_EDC" - The USA Contiguous Lambert Azmuthal Equal Area (USGS-EDC) Projection. (https://spatialreference.org/ref/sr-org/6903/)
"AUS_MERC" - The Ahughes Australian Mercator (https://spatialreference.org/ref/epsg/3514/)
"ARG_BI" - The Argentina Bicontinental - South Pole Stereographic (https://spatialreference.org/ref/sr-org/7098/) 

```{r}
#creating different projection options
USGS_EDC <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"
AUS_MERC <- "+proj=utm +zone=54 +south +ellps=aust_SA +units=m +no_defs "
ARG_BI <- "+proj=stere +lat_0=-90 +lat_ts=-90 +lon_0=-63 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
```

```{r}
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = USGS_EDC) +
  theme_bw()
```
```{r}
#Inspired by the projections Kristy H. shared with us in class.
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = AUS_MERC) +
  theme_bw()
```
```{r}
ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = ARG_BI) +
  theme_bw()
```

```{r}
states_transformed <- US_states %>%
  st_transform(crs = USGS_EDC)

ggplot(states_transformed) +
  geom_sf() +
  theme_bw()
```
The variable I chose to explore was Median Household Income by State. The data originally comes from the 2018 ACS 1-Year Survey. I used a .csv file created by another Github user in my analysis.

```{r}
#Inspired by Ana Z's Canvas discussion post
urlfile="https://raw.githubusercontent.com/Ajstanley89/State-of-Healthcare/master/assets/data/data.csv"
income_data <- read.csv(url(urlfile))
head(income_data)
income_states <-states_transformed %>%
  left_join(income_data, by = c("postal" = "abbr")) %>%
  select(name, state, income, poverty, age, incomeMoe, id)
```

```{r}
#legend.position code inspired by Ignacio L.
ggplot(income_states, aes(fill = income)) +
  geom_sf(color = "lightblue") +
  scale_fill_viridis_c(
    guide = "colourbar", aesthetics = "fill", 
    name = "Thousands of Dollars",
    breaks = seq(30000, 80000, by = 5000),
    labels = formatC(seq(30000, 80000, by = 5000), 
                     big.mark = ",", format = "f", digits = 0)) +
  ggtitle("                 Median Household Income by State") +
  theme_map() +
  theme(legend.background = element_blank(),
        legend.position = "left")
```

```{r}
income_centeroids <- income_states %>%
  st_centroid()

ggplot(states_transformed) +
  geom_sf(fill = NA, color = "black") +
  geom_sf(data = income_centeroids, 
          aes(size = income),
          alpha = 0.2, color = "chartreuse3") +
  scale_size_continuous(name = "Median Income by State",
    breaks = seq(100000, 500000, by = 100000),
    labels = formatC(seq(100000, 500000, by = 100000), 
                     big.mark = ",", format = "f", digits = 0),
    range = c(0, 20)) +
    ggtitle(label = "Median Household Income by State in Thousands of Dollars", "Minimum (Mississippi): $39680, Maximum (Maryland): $73971") +
    theme(plot.title = element_text(hjust = 0.5)) +
  theme_void() 
```
```{r}
income_cartogram_ncont <- income_states %>%
  cartogram_ncont("age")

ggplot(income_cartogram_ncont, aes(fill = income)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Median Household Income by State in Thousands of Dollars",
    breaks = seq(100000, 500000, by = 100000),
    labels = formatC(seq(100000, 500000, by = 100000), 
                     big.mark = ",", format = "f", digits = 0)) +
   ggtitle(label = "Median Household Income by State in Thousands of Dollars", "Minimum (Mississippi): $39680, Maximum (Maryland): $73971") +
  theme_map() +
  theme(legend.background = element_blank(), legend.position = "left")
```
```{r}
income_cartogram_cont <- income_states %>%
  cartogram_cont("income")

ggplot(income_cartogram_cont, aes(fill = income)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Median Household Income by State in Thousands of Dollars",
    breaks = seq(100000, 500000, by = 100000),
    labels = formatC(seq(100000, 500000, by = 100000), 
                     big.mark = ",", format = "f", digits = 0)) +
   ggtitle(label = "Median Household Income by State in Thousands of Dollars", "Minimum (Mississippi): $39680, Maximum (Maryland): $73971") +
  theme_map() +
  theme(legend.background = element_blank())
```

