---
title: "Assignment 1"
author: "Gianina Yumul"
date: "9/14/2020"
output: html_document
---
```{r}
library(tidycensus)
library(tidyverse)
library(magrittr)
library(ggthemes)
library(ggplot2)
library(wesanderson)
```

Variables pulled from ACS 5-year community data:

1. Majority racial/ethnic identity (categorical)
2. Citizenship status (categorical)
3. Median age (continuous)
4. Average household size (continuous)
5. Percent of households who live in rented homes (continuous)
6. Population of tract (continuous)

```{r}
<<<<<<< HEAD
vars <- c(tot_hh = "B08137_001E",
           tot_rent_hh = "B08137_003E",
           avg_hh_size = "B25010_001E",
           tot_pop = "B02001_001E",
           nh_white_pop = "B02001_002E",
           nh_black_pop = "B02001_003E",
           nh_asian_pop = "B02001_005E",
           hispanic_pop = "B01001I_001E",
           tot_cs_all = "B05001_001E",
           tot_cs_bornus = "B05001_002E",
           tot_cs_bornpr = "B05001_003E",
           tot_cs_nat = "B05001_005E",
           tot_cs_none = "B05001_006E",
           med_age = "B01002_001E")
=======
vars <- c(tot_hh = "H001001",
           tot_rural_hh = "H002005",
           tot_rent_hh = "H004004",
           avg_hh_size = "H012001",
           tot_pop = "P003001",
           nh_white_pop = "P005003",
           nh_black_pop = "P005004",
           nh_asian_pop = "P005006",
           hispanic_pop = "P005010",
           
           med_age = "P013001")
>>>>>>> 9cf80514130c5dc0906a62588407d9c70ac0e5c3
```

I obtained data for tracts in Alameda County, CA.

```{r}
AL_tracts <- get_acs(geography = "tract", variables = vars,
                           state = "CA", county = "Alameda",
                           output = "wide") %>%
  mutate(pct_rental = tot_rent_hh / tot_hh,
         pct_bornus = tot_cs_bornus / tot_cs_all,
         pct_bornpr = tot_cs_bornpr / tot_cs_all,
         pct_nat = tot_cs_nat / tot_cs_all,
         pct_none = tot_cs_none / tot_cs_all,
         majority_race_eth = case_when(
           nh_white_pop / tot_pop > 0.5 ~ "nh_white",
           nh_black_pop / tot_pop > 0.5 ~ "nh_black",
           nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
           hispanic_pop / tot_pop > 0.5 ~ "hispanic",
           (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /tot_pop < 0.5 ~ "other",
           TRUE ~ "no_majority")) %>%
  filter(tot_pop > 0, tot_hh > 0) %>%
  select(GEOID, majority_race_eth, 
         pct_rental, avg_hh_size, tot_pop, med_age, pct_bornus, pct_bornpr, pct_nat, pct_none) 
```

This is graph #1. This graph shows average household size as a function of the percent of households in rented homes in each tract. The fill shows the majority race/ethnicity of each tract along with a count for the number of observations represented by each dot.

I don't think this is a particularly good graph. It is quite hard to read and it takes time to adjust to. After playing with geom_bin2d function, I would like to keep it in my future 'toolbox' and think about scenarios in which using both the outline and fill of a point in a scatterplot to communicate something might be appropriate.

```{r}
ggplot(AL_tracts, 
       aes(x = pct_rental,
           y = avg_hh_size,
           color = majority_race_eth)) +
  geom_bin2d() +
  scale_x_continuous(name = "Percent of households in rented homes",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Average Household Size",
                     breaks = seq(0, 6, by = 1)) +
  scale_color_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
  theme_tufte() 
```

This is graph #2. This graph shows median age within a tract as a function of the percent of households in mortgaged/owned homes, along with trendlines for each majority race/ethnicity. Conceptually, this is very similar to my first chart. One new aspect I wanted to explore in this chart was the possibility of using gradations of a single color to communicate trends within each of these subgroups.

```{r}
ggplot(AL_tracts, 
       aes(x = 1 - pct_rental,
           y = med_age,
           color = majority_race_eth,
           fill = majority_race_eth)) +
  geom_point(alpha  = 1/10) +
  stat_smooth(method = "lm") +
  scale_x_continuous(name = "Percent of households in mortgaged/owned homes",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_brewer(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_fill_brewer(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
  theme_classic()
```

This is graph #3. I believe this graph shows a lack of strong correlation between a tract's population and the median age of the tract. The outliers interest me in this representation; in particular, I was curious about the six tracts where the median age was less than 25 and the tract populations varied between 1250 and 9000.  

```{r}
ggplot(AL_tracts, 
       aes(x = med_age,
           y = tot_pop)) +
  geom_point(color = wes_palette(n = 1, name = "GrandBudapest2")) +
  stat_smooth(color = wes_palette(n = 1, name = "GrandBudapest2"), linetype = 2, size = 0.5) +
  scale_x_continuous(name = "Median age of tract") +
  scale_y_continuous(name = "Tract population") +
  ggtitle("Median Age of ACS Tracts", subtitle = "Alameda County, CA") +
  theme_solarized() 
```

This is graph #4. In this graph, I show average household size as a function of the percentage of US-born citizens within tracts. The chart shows that, in general, the average household size decreases as the percentage of US-born citizens within a tract increases.

```{r}
ggplot(AL_tracts,
       aes(x = pct_bornus,
           y = avg_hh_size)) +
  geom_jitter(color = wes_palette(n = 1, name = "Moonrise2")) +
  scale_x_continuous(name = "Percent of US-born Citizens",
                     labels = scales::percent) +
  scale_y_continuous(name = "Average Household Size",
                     breaks = seq(0, 6, by = 1)) +
  stat_smooth(method = "lm") +
  ggtitle("Household Size vs. Percent of US-born Citizens", subtitle = "Alameda County, CA") +
  theme_tufte() 
```

This is graph #5, taking full advantage of the Tufte themes. This graph shows the median age of households in tracts with a majority race/ethnicity. The 50th percentile for each majority race/ethnicity category is prominent and is easy to comparee across categories.

```{r}
ggplot(AL_tracts, 
       aes(x = majority_race_eth,
           y = med_age)) +
  geom_tufteboxplot() +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Median age within tract",
                     breaks = seq(0, 100, by = 10)) +
  ggtitle("Median Age of Households in Tracts \nwith a Majority Race/Ethnicity", subtitle = "Alameda County, CA") +
  theme_tufte() 
```

This is graph #6, which shows the percentage of naturalized citizens in tracts with a majority race/ethnicity. One thing I found interesting about this representation is that it illustrates that the non-Hispanic Asian category had the widest spread and highest maximum value of naturalized citizens within a tract, with one tract nearing 50% naturalized citizens within its population.

```{r}
ggplot(AL_tracts, 
       aes(x = majority_race_eth,
<<<<<<< HEAD
           y = pct_nat,
=======
           y = avg_hh_size,
>>>>>>> 9cf80514130c5dc0906a62588407d9c70ac0e5c3
           color = majority_race_eth)) +
  geom_point(position = "jitter",
             size = 1.5,
             alpha = 0.5) +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
<<<<<<< HEAD
  scale_y_continuous(name = "Percentage of Naturalized Citizens",
                     labels = scales::percent) +
=======
  scale_y_continuous(name = "Average household size",
                     breaks = seq(0, 5, by = 1)) +
>>>>>>> 9cf80514130c5dc0906a62588407d9c70ac0e5c3
  theme_bw() +
  ggtitle("Percentage of Naturalized Citizens \nin Tracts with a Majority Race/Ethnicity", subtitle = "Alameda County, CA") +
  theme(legend.position = "none")
```

This is graph #7, which shows the percentage of homes mortgaged/owned in tracts with a majority race/ethnic group. Like the Tufte-inspired graph above, I thought these boxplots were clear and easy to compare across groups.
```{r}
ggplot(AL_tracts, 
       aes(x = majority_race_eth,
           y = 1 - pct_rental,
           color = majority_race_eth)) +
  geom_boxplot(fill='#FFA07A', color="darkred") +
  scale_x_discrete(name = "MAJORITY RACE/ETHNICITY",
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nblack",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  scale_y_continuous(name = "Percent of homes mortgaged/owned",
                     labels = scales::percent) +
  theme_economist() +
  ggtitle("Percentage of Homes Mortgaged/Owned \nin Tracts with a Majority Race/Ethnic Group", subtitle = "Alameda County, CA")
  theme(legend.position = "none")
```

This is graph #8. I used a violin plot to compare average household size between tracts with a majority race/ethnic group.

```{r}
ggplot(AL_tracts, 
       aes(x = majority_race_eth,
           y = avg_hh_size,
           color = majority_race_eth)) +
  geom_violin() +
  scale_color_colorblind() +
  coord_flip() +
  stat_summary(fun.y=mean, geom = "point", shape = 23, size = 2) +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic black",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Average household size within tract",
                     breaks = seq(0, 4, by = 1)) +
  theme_bw() +
  theme(legend.position = "none")
```


This is graph #9, a stacked bar graph that shows the relative percentages of majority race/ethnic groups with average household sizes of between 1 and 4.
```{r}
ggplot(AL_tracts, aes(x = ceiling(avg_hh_size), fill = majority_race_eth)) +
  geom_bar(position = "fill") +
  scale_color_colorblind() +
  scale_y_continuous(name = "percent of tracts",
                     labels = scales::percent) +
  scale_x_discrete(name = "Household Size",
                    labels = c("1","2","3","4")) +
  scale_fill_discrete(name = element_blank(),
                        labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nBlack",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  theme_bw() 
```


This is graph #10. This scatter plot shows the percentage of homes mortgaged/owned as a function of the percentage of non-citizens within a tract.
```{r}
ggplot(AL_tracts,
       aes(x = pct_none,
           y = pct_rental, color = majority_race_eth)) +
  geom_jitter(color = wes_palette(n = 1, name = "Moonrise2")) +
  scale_x_continuous(name = "Percent of Non-US Citizens",
                     labels = scales::percent) +
  scale_y_continuous(name = "Percent of Homes Mortgaged/Owned",
                     labels = scales::percent) +
  stat_smooth(method = "lm") +
  ggtitle("Household Size vs. Percent of Non-US Citizens", subtitle = "Alameda County, CA") +
  theme_tufte() 
  
```

